Signatures and Induction Principles for Higher Inductive-Inductive Types

### 2019-02-04 paper.tex
stuff

%% read in additional TeX-packages or personal macros here:
%% e.g. \usepackage{tikz}
%% \usepackage[dvipsnames]{xcolor}
\usepackage{microtype}
\usepackage{proof}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\bibliographystyle{plainurl}
\usepackage{hyperref}

@@

(5) There is a different equality type former, which can be used to express
paths between type constructors. To our knowledge, this has not been considered
so far in HIT literature, although Cartmell considered equations between sorts
in generalized algebraic theories \cite{gat}. For an example, paths between type
constructors allow a compact definition of integers:
\begin{alignat*}{5}
  & \boldsymbol{\cdot},\,\,Int : U,\,\,zero : \underline{Int},\,p : Int =_\U Int
\end{alignat*}
Here, successor and predecessor functions can be recovered from the $p : Int
=_\U Int$ equality, by transporting an $Int$ along $p$ or $p^{-1}$. This
signature could be unfolded to a larger one, by replacing $p : Int =_\U Int$
with an explicit $suc : Int \ra \underline{Int}$ constructor and additional
constructors expressing that $suc$ is an equivalence. The definition of $Int$
with a successor equivalence was previously suggested by Altenkirch, Pinyo
[TODO] and Cavallo [TODO]. Our signature can be viewed as a shorthand for the
unfolded version.

>>>>

%% read in additional TeX-packages or personal macros here:
%% e.g. \usepackage{tikz}
%% \usepackage[dvipsnames]{xcolor}
\usepackage{microtype}
\usepackage{proof}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\bibliographystyle{plainurl}
\usepackage{hyperref}
\usepackage{url}

@@

(5) There is a different equality type former, which can be used to express
paths between type constructors. To our knowledge, this has not been considered
so far in HIT literature, although Cartmell considered equations between sorts
in generalized algebraic theories \cite{gat}. For an example, paths between type
constructors allow a compact definition of integers:
\begin{alignat*}{5}
  & \boldsymbol{\cdot},\,\,Int : U,\,\,zero : \underline{Int},\,p : Int =_\U Int
\end{alignat*}
Here, successor and predecessor functions can be recovered from the $p : Int
=_\U Int$ equality, by transporting an $Int$ along $p$ or $p^{-1}$. This
signature could be unfolded to a larger one, by replacing $p : Int =_\U Int$
with an explicit $suc : Int \ra \underline{Int}$ constructor and additional
constructors expressing that $suc$ is an equivalence. The definition of $Int$
with a successor equivalence was previously suggested by Altenkirch and Pinyo
\cite{pinyo2018integers} and Cavallo and M\"ortberg \cite{cavallointegers}. Our
signature can be viewed as a shorthand for the unfolded version.

For another example, we may have a HIIT definition for a type theory where
Russell-style universes are compactly specified with an equality:
\begin{alignat*}{5}
  & ...,\,\, russell : (\Gamma : Con)\ra Tm\,\Gamma\,U =_\U Ty\,\Gamma,\,\, ...
\end{alignat*}
